---
title: "STA 235H - Review Session I"
subtitle: "Fall 2022"
author: "McCombs School of Business, UT Austin"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js"]
    includes:
      in_header: ["header.html"]
---

<!-- <script type="text/javascript"> -->
<!-- MathJax.Hub.Config({ -->
<!--   "HTML-CSS": { -->
<!--     preferredFont: null, -->
<!--     webFont:  "Neo-Euler" -->
<!--   } -->
<!-- }); -->
<!-- </script> -->

```{r setup, include=FALSE, warning=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE, message=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Yanone Kaffeesatz",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed","200","300","400","500","600")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Jost"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "115%",
  colors = c(
    lightgrey = "#C0C0C0",
    red = "#f34213",
    purple = "#900DA4",
    darkpurple = "#61077a",
    darkorange = "#db5f12",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "none"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".tiny" = list("font-size" = "50%"),
    ".large" = list("font-size" = "150%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Jost",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
   ".bottom2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "90%",
                      "left" = "10%",
                      "-ms-transform" = "translate(-10%, -90%)",
                      "transform" = "translate(-10%, -90%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1s,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2s,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3s,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4s,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5s,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6s,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7s,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25"),
   
    ".box-1trans,.box-2trans,.box-3trans,.box-4trans,.box-5trans,.box-6trans,.box-7trans" = list("background-color" = "#FFFFFF",
                                                                                                 "font-family" = "Yanone Kaffeesatz",
                                                                                                 "border-radius" = "15px",
                                                                                                 "margin" = "0em auto",
                                                                                                 "overflow" = "hidden",
                                                                                                 "padding" = "0.4em 0.4em",
                                                                                                 "font-weight" = "600",
                                                                                                 "font-size" = "31px",
                                                                                                 "display" = "table",
                                                                                                 "text-align" = "center",
                                                                                                 "border-width" = "thick",
                                                                                                 "color" = "#333f48"),
   
       ".box-1Trans,.box-2Trans,.box-3Trans,.box-4Trans,.box-5Trans,.box-6Trans,.box-7Trans" = list("background-color" = "#FFFFFF",
                                                                                                 "font-family" = "Yanone Kaffeesatz",
                                                                                                 "border-radius" = "15px",
                                                                                                 "margin" = "0em auto",
                                                                                                 "overflow" = "hidden",
                                                                                                 "padding" = "0.4em 0.4em",
                                                                                                 "font-weight" = "600",
                                                                                                 "font-size" = "51px",
                                                                                                 "display" = "table",
                                                                                                 "text-align" = "center",
                                                                                                 "border-width" = "thick",
                                                                                                 "color" = "#333f48"),
   
       ".box-1transl,.box-2transl,.box-3transl,.box-4transl,.box-5transl,.box-6transl,.box-7transl" = list("background-color" = "#FFFFFF",
                                                                                                           "font-family" = "Yanone Kaffeesatz",
                                                                                                           "border-radius" = "15px",
                                                                                                           "left" = "0px",
                                                                                                           "overflow" = "hidden",
                                                                                                           "padding" = "0.4em 0.4em",
                                                                                                           "font-weight" = "600",
                                                                                                           "font-size" = "31px",
                                                                                                           "display" = "table",
                                                                                                           "text-align" = "left",
                                                                                                           "border-width" = "thick",
                                                                                                           "color" = "#333f48"),
   
   
    ".box-1trans,.box-1transl,.box-1Trans" = list("border"="5px solid rgba(13, 8, 135,1)"),
    ".box-2trans,.box-2transl,.box-2Trans" = list("border"="5px solid rgba(86, 1, 164,1)"),
    ".box-3trans,.box-3transl,.box-3Trans" = list("border"="5px solid rgba(144, 13, 164,1)"),
    ".box-4trans,.box-4transl,.box-4Trans" = list("border"="5px solid rgba(191, 57, 132,1)"),
    ".box-5trans,.box-5transl,.box-5Trans" = list("border"="5px solid rgba(225, 100, 98,1)"),
    ".box-6trans,.box-6transl,.box-6Trans" = list("border"="5px solid rgba(248, 148, 65,1)"),
    ".box-7trans,.box-7transl,.box-7Trans" = list("border"="5px solid rgba(252, 206, 37,1)"),

   
   ".box-1t,.box-1tL,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.3)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-2t,.box-2tL,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.3)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-3t,.box-3tL,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.3)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-4t,.box-4tL,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.3)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-5t,.box-5tL,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.3)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-6t,.box-6tL,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.3)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-7t,.box-7tL,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.3)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
   ".box-7tL, .box-6tL, .box-5tL, .box-4tL, .box-3tL, .box-2tL, .box-1tL" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "50px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7s, .box-6s, .box-5s, .box-4s, .box-3s, .box-2s, .box-1s" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.2em 0.2em",
                                                                      "font-weight" = "400",
                                                                      "font-size" = "100%",
                                                                      "display" = "inline",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".medium" = list("font-size" = "35px"),
   
   ".medium30" = list("font-size" = "30px"),
   
   ".jost-normal" = list("font-family" = "Jost",
                         "font-size" = "31px"),
   
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "67%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "27%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")


  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```
```{css, echo = FALSE}

.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}

.tiny .remark-code { /*Change made here*/
  font-size: 90% !important;
}
```

```{r setup2, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(showtext)
library(xaringanExtra)

#xaringanExtra::use_scribble()

htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

```{r fonts, message=FALSE, echo=FALSE, warning=FALSE}
font.add.google("Fira Sans Condensed", "Fira Sans Condensed")
font.add.google("Fira Sans", "Fira Sans")
```

# Structure

- We will review **.darkorange[different types of regressions]**

--

- I will provide a general framework of **.darkorange[how to think about these regressions]**

--

- We will do one exercise **.darkorange[together]**

--

.box-7Trans[Lightning round!]

---
<br>
<br>

.box-2Trans[Participate!]

.box-2trans[Even if you make a mistake, everyone can learn from that]

--

.box-4Trans[Ask questions!]

.box-4trans[You are here on a Friday afternoon.. take advantage of it :)]

---
# Typical regression: Continuous variables
<br>
<br>
<br>

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \varepsilon$$

- $Y$ is a continuous variable

- $X$ is a continuous variable

--

.center[
*A one-unit increase in $X_1$ is associated with an average increase in $Y$ of $\beta_1$ units, holding $X_2$ constant*
]

---
# Typical regression: Example

Let's analyze what factors associate to someone's grade:

$$Grade = \beta_0 + \beta_1 HrsAssign + \beta_2PreGrade + \beta_3ClassAtt + \varepsilon$$

Where:

- **Grade**: Grade in an assignment (out of 100 points).
- **HrsAssign**: Hours dedicated to the assignment.
- **PreGrade**: Grade in the previous assignment (out of 100 points).
- **ClassAtt**: Whether someone attended the last class (1) or not (0).


.box-5trans[Interpret the coefficient for Hours]

---
# Typical regression: Example

.small[
```{r, echo=FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(truncnorm)

set.seed(100)

n = 200

hours = rtruncnorm(n, a=0, b=20, mean = 5, sd = 4)

hours = round(hours, 2)

pregrade = rtruncnorm(n, a=50, b=100, mean = 92, sd = 6)

classatt = sample(c(0,1), n, replace = TRUE, prob = c(0.25,0.75))

grade = 1.4 + 3.1*hours + 0.67*pregrade + 1.1*classatt + rnorm(n, mean = 0, sd = 1.5)  

#summary(grade)

grade_curve = (100 - 60)/(118.75 - 54.79)*(grade - 118.75) + 100

#summary(grade_curve)

d <- data.frame(hours = hours,
                pregrade = pregrade,
                classatt = classatt,
                grade = grade_curve)


#summary(d$grade)
```
```{r}
summary(lm(grade ~ hours + pregrade + classatt, data = d))
``` 
]

---
# Typical regression: Binary covariate
<br>
<br>
<br>

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \varepsilon$$

- $Y$ is a continuous variable

- $X$ is a binary variable

--

.center[
<i>The event of $X_1=1$ is associated with an average increase in $Y$ of $\beta_1$ units compared to $X_1=0$, holding $X_2$ constant</i>
]

---
# Typical regression: Example

Let's analyze what factors associate to someone's grade:

$$Grade = \beta_0 + \beta_1 HrsAssign + \beta_2PreGrade + \beta_3ClassAtt + \varepsilon$$

Where:

- **Grade**: Grade in an assignment (out of 100 points).
- **HrsAssign**: Hours dedicated to the assignment.
- **PreGrade**: Grade in the previous assignment (out of 100 points).
- **ClassAtt**: Whether someone attended the last class (1) or not (0).


.box-5trans[Interpret the coefficient for Class Attendance]

---
# Typical regression: Example

- Interpret the coefficient for *ClassAtt*:

.small[
```{r}
summary(lm(grade ~ hours + pregrade + classatt, data = d))
``` 
]

---
# Regressions with logarithms: Log-level

<br>
<br>
<br>

$$\log(Y) = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \varepsilon$$

- The outcome variable is in a logarithm, $\log(Y)$

- $X$ is a continuous variable, not in a logarithm.

--

.center[
*A one-unit increase in $X_1$ is associated with an average increase in $Y$ of $\beta_1\times100$ percent, holding $X_2$ constant*
]

---
# Regressions with logarithms: Example

Let's analyze what characteristics are associated with sales on Etsy:

$$\log(Sales) = \beta_0 + \beta_1 Price + \beta_2AvgRating + \beta_3NReviews + \varepsilon$$

Where:

- **log(Sales)**: Weekly sales of a product ($), in a logarithm.
- **Price**: Price of the product ($)
- **AvgRating**: Average rating of the product (in a scale fo 1-5).
- **NReviews**: Number of reviews of a product.


.box-5trans[Interpret the coefficient for Number of Reviews]

---
# Regressions with logarithms: Example

- Interpret the coefficient for *NReviews*:

.small[
```{r, echo = FALSE, message = FALSE, warning=FALSE}
set.seed(100)

n = 300

price = rtruncnorm(n, a=10, b=200, mean = 35, sd = 10)

avgrating = rtruncnorm(n, a=1, b=5, mean = 4, sd = 1)

avgrating = round(avgrating, 1)

nreviews = rtruncnorm(n, a=0, b=1200, mean = 45, sd = 20)

nreviews = round(nreviews, 0)

sales = 100 + price*1.05 + avgrating*50 + nreviews*30 + rnorm(n, mean = 0, sd = 79)  

#summary(sales)


d <- data.frame(price = price,
                avgrating = avgrating,
                nreviews = nreviews,
                sales = sales)

```

```{r}
summary(lm(log(sales) ~ price + avgrating + nreviews, data = d))
```
]

---
# Regressions with logarithms: Your turn

- Interpret the coefficient for *Price*:

.small[
```{r}
summary(lm(log(sales) ~ price + avgrating + nreviews, data = d))
```
]

---
# Regressions with Interactions
<br>
<br>
<br>

$$Y = \beta_0 + \beta_1 X_1 + \beta_2D +  \beta_3X_1\times D + \beta_4 X_2 + \varepsilon$$

- $Y$ is a continuous variable

- $X$ is a continuous variable

- $D$ is a binary variable

--

.center[
*A one-unit increase in $X_1$ is associated with an average increase in $Y$ of $\beta_1$ units for $D=0$, holding $X_2$ constant*
]

--

.center[
*A one-unit increase in $X_1$ is associated with an average increase in $Y$ of $\beta_1 + \beta_3$ units for $D=1$, holding $X_2$ constant*
]


---
# Why?




---
# Regressions with Interactions: Example

Let's analyze what factors associate to someone's grade:

$$Grade = \beta_0 + \beta_1PreGrade + \beta_2 HrsAssign + \beta_3ClassAtt + \beta_4HrsAssign \times ClassAtt + \varepsilon$$

Where:

- **Grade**: Grade in an assignment (out of 100 points).
- **HrsAssign**: Hours dedicated to the assignment.
- **PreGrade**: Grade in the previous assignment (out of 100 points).
- **ClassAtt**: Whether someone attended the last class (1) or not (0).


.box-5trans[Interpret the association between Hours and Grade for students that attended the last class and those that did not]

---
# Regressions with Interactions: Example

- Interpret the association the assicuation between *hours* and *grade* for the two groups.

.small[
```{r, echo = FALSE, warning = FALSE, message=FALSE}
set.seed(100)

n = 200

grade = 1.4 + 3.1*hours + 0.67*pregrade + 1.1*hours*classatt + rnorm(n, mean = 0, sd = 1.5)  

#summary(grade)

grade_curve = (100 - 60)/(132.71 - 56.4)*(grade - 132.71) + 100

#summary(grade_curve)

d <- data.frame(hours = hours,
                pregrade = pregrade,
                classatt = classatt,
                grade = grade_curve)


#summary(d$grade)
```

```{r}
summary(lm(grade ~ pregrade + hours*classatt, data = d))
```
]
---
# Quadratic Regressions
<br>
<br>
<br>

$$Y = \beta_0 + \beta_1 X_1 + \beta_2X_1^2 + \beta_3 X_2 + \varepsilon$$

- $Y$ is a continuous variable

- $X$ is a continuous variable and includes a quadratic term

--

.center[
*Increasing $X_1$ in one unit, from $x_0$ to $x_0+1$, is associated with an average increase in Y of $2\beta_2 x_0 + \beta_1$ units, holding $X_2$ constant*
]

---
# Quadratic Regressions: Example

We think there is a quadratic association between price and sales, so we fit the following model:

$$Sales = \beta_0 + \beta_1 Price + \beta_2Price^2  + \beta_3AvgRating + \beta_4NReviews + \varepsilon$$

Where:

- **Sales**: Weekly sales of a product ($)
- **Price**: Price of the product ($)
- **AvgRating**: Average rating of the product (in a scale fo 1-5).
- **NReviews**: Number of reviews of a product.


.box-5trans[Interpret the association between Sales and Price, for an increase in Price from &dollar;15 to &dollar;16]

---
# Quadratic Regressions: Example

- Interpret the association between *sales* and *price* (for an increase in price from 15 to 16 dollars)

.small[
``` {r, echo = FALSE, message=FALSE, warning=FALSE}

set.seed(100)

n = 300

price = rtruncnorm(n, a=10, b=200, mean = 35, sd = 10)

avgrating = rtruncnorm(n, a=1, b=5, mean = 4, sd = 1)

avgrating = round(avgrating, 1)

nreviews = rtruncnorm(n, a=0, b=1200, mean = 45, sd = 20)

nreviews = round(nreviews, 0)

sales = 100 - price*10.05 + price^2*(2.5) + avgrating*50 + nreviews*30 + rnorm(n, mean = 0, sd = 79)  

#summary(sales)


d <- data.frame(price = price,
                avgrating = avgrating,
                nreviews = nreviews,
                sales = sales)


#summary(d$sales)
```

```{r}
summary(lm(sales ~ price + I(price^2) + avgrating + nreviews, data = d))
``` 
]

---
# Linear Probability Model

<br>
<br>
<br>

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \varepsilon$$

- $Y$ is a binary variable

- $X$ is a continuous variable

--

.center[
*Increasing $X_1$ in one unit is associated with an average increase in the probability of $Y=1$ of $\beta_1$, holding $X_2$ constant*
]


.center[
*Increasing $X_1$ in one unit is associated with an average increase in the probability of $Y=1$ of $\beta_1\times100$ percentage points, holding $X_2$ constant*
]

---
# Linear Probability Model: Example

Let's analyze what factors associate to getting an A in an assignment:

$$Grade = \beta_0 + \beta_1 HrsAssign + \beta_2PreGrade + \beta_3ClassAtt + \varepsilon$$

Where:

- **GradeA**: Binary variable if the grade in the assignment is A (1) or not (0).
- **HrsAssign**: Hours dedicated to the assignment.
- **PreGrade**: Grade in the previous assignment (out of 100 points).
- **ClassAtt**: Whether someone attended the last class (1) or not (0).


.box-5trans[Interpret the coefficient for Hours]

---
# Linear Probability Model: Example

- Interpret the coefficient for *Hours*.

.small[
```{r, echo=FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(truncnorm)

set.seed(100)

n = 200

hours = rtruncnorm(n, a=0, b=20, mean = 5, sd = 4)

hours = round(hours, 2)

pregrade = rtruncnorm(n, a=50, b=100, mean = 92, sd = 6)

classatt = sample(c(0,1), n, replace = TRUE, prob = c(0.25,0.75))

grade = 1.4 + 3.1*hours + 0.67*pregrade + 1.1*classatt + rnorm(n, mean = 0, sd = 1.5)  

#summary(grade)

grade_curve = (100 - 60)/(118.75 - 54.79)*(grade - 118.75) + 100

#summary(grade_curve)

d <- data.frame(hours = hours,
                pregrade = pregrade,
                classatt = classatt,
                grade = grade_curve)

d <- d %>% mutate(gradeA = ifelse(grade>90, 1, 0))

#summary(d$grade)
```
```{r}
summary(lm(gradeA ~ hours + pregrade + classatt, data = d))
``` 
]

---
# Putting everything together

Let's interpret what is the return to experience (i.e. association between income and experience) for men and women:

$$Income = \beta_0 + \beta_1education + \beta_2exp + \beta_3exp^2 + \beta_4 female + \beta_5exp\times female + \beta_6exp^2\times female + \varepsilon$$

---
# Putting everything together

.small[
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
## Income gender

set.seed(100)

n = 1200

age = rtruncnorm(n, a=18, b=65, mean = 45, sd = 5)

education = rtruncnorm(n, a=5, b=22, mean = 13, sd = 4)

female = sample(c(0,1), n, replace = TRUE)

experience = age - education + 5 - 4*female

income = 15674 + education*1997 + experience*1345 + experience^2*(-9) - 2400*female - 437*female*experience - 20.2*female*experience^2 + rnorm(n, mean = 0, sd = 5300)  

#summary(income)


d <- data.frame(education = education,
                experience = experience,
                female = female,
                income = income)

```
```{r}
summary(lm(income ~ education + experience*female + I(experience^2)*female, data = d))
``` 
]

---

