---
title: "STA 235H - Heteroskedasticity"
subtitle: "Fall 2022"
author: "McCombs School of Business, UT Austin"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css]    
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js"]
    includes:
      in_header: ["header.html"]
---

<!-- <script type="text/javascript"> -->
<!-- MathJax.Hub.Config({ -->
<!--   "HTML-CSS": { -->
<!--     preferredFont: null, -->
<!--     webFont:  "Neo-Euler" -->
<!--   } -->
<!-- }); -->
<!-- </script> -->

```{r setup, include=FALSE, warning=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE, message=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Yanone Kaffeesatz",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed","200","300","400","500","600")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Jost"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "115%",
  colors = c(
    lightgrey = "#C0C0C0",
    red = "#f34213",
    purple = "#900DA4",
    darkpurple = "#61077a",
    darkorange = "#db5f12",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "none"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".tiny" = list("font-size" = "50%"),
    ".large" = list("font-size" = "150%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Jost",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
   ".bottom2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "90%",
                      "left" = "10%",
                      "-ms-transform" = "translate(-10%, -90%)",
                      "transform" = "translate(-10%, -90%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1s,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2s,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3s,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4s,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5s,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6s,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7s,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25"),
   
    ".box-1trans,.box-2trans,.box-3trans,.box-4trans,.box-5trans,.box-6trans,.box-7trans" = list("background-color" = "#FFFFFF",
                                                                                                 "font-family" = "Yanone Kaffeesatz",
                                                                                                 "border-radius" = "15px",
                                                                                                 "margin" = "0em auto",
                                                                                                 "overflow" = "hidden",
                                                                                                 "padding" = "0.4em 0.4em",
                                                                                                 "font-weight" = "600",
                                                                                                 "font-size" = "31px",
                                                                                                 "display" = "table",
                                                                                                 "text-align" = "center",
                                                                                                 "border-width" = "thick",
                                                                                                 "color" = "#333f48"),
   
       ".box-1Trans,.box-2Trans,.box-3Trans,.box-4Trans,.box-5Trans,.box-6Trans,.box-7Trans" = list("background-color" = "#FFFFFF",
                                                                                                 "font-family" = "Yanone Kaffeesatz",
                                                                                                 "border-radius" = "15px",
                                                                                                 "margin" = "0em auto",
                                                                                                 "overflow" = "hidden",
                                                                                                 "padding" = "0.4em 0.4em",
                                                                                                 "font-weight" = "600",
                                                                                                 "font-size" = "51px",
                                                                                                 "display" = "table",
                                                                                                 "text-align" = "center",
                                                                                                 "border-width" = "thick",
                                                                                                 "color" = "#333f48"),
   
       ".box-1transl,.box-2transl,.box-3transl,.box-4transl,.box-5transl,.box-6transl,.box-7transl" = list("background-color" = "#FFFFFF",
                                                                                                           "font-family" = "Yanone Kaffeesatz",
                                                                                                           "border-radius" = "15px",
                                                                                                           "left" = "0px",
                                                                                                           "overflow" = "hidden",
                                                                                                           "padding" = "0.4em 0.4em",
                                                                                                           "font-weight" = "600",
                                                                                                           "font-size" = "31px",
                                                                                                           "display" = "table",
                                                                                                           "text-align" = "left",
                                                                                                           "border-width" = "thick",
                                                                                                           "color" = "#333f48"),
   
   
    ".box-1trans,.box-1transl,.box-1Trans" = list("border"="5px solid rgba(13, 8, 135,1)"),
    ".box-2trans,.box-2transl,.box-2Trans" = list("border"="5px solid rgba(86, 1, 164,1)"),
    ".box-3trans,.box-3transl,.box-3Trans" = list("border"="5px solid rgba(144, 13, 164,1)"),
    ".box-4trans,.box-4transl,.box-4Trans" = list("border"="5px solid rgba(191, 57, 132,1)"),
    ".box-5trans,.box-5transl,.box-5Trans" = list("border"="5px solid rgba(225, 100, 98,1)"),
    ".box-6trans,.box-6transl,.box-6Trans" = list("border"="5px solid rgba(248, 148, 65,1)"),
    ".box-7trans,.box-7transl,.box-7Trans" = list("border"="5px solid rgba(252, 206, 37,1)"),

   
   ".box-1t,.box-1tL,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.3)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-2t,.box-2tL,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.3)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-3t,.box-3tL,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.3)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-4t,.box-4tL,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.3)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-5t,.box-5tL,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.3)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-6t,.box-6tL,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.3)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
    ".box-7t,.box-7tL,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.3)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
   ".box-7tL, .box-6tL, .box-5tL, .box-4tL, .box-3tL, .box-2tL, .box-1tL" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "50px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "border-radius" = "15px"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7s, .box-6s, .box-5s, .box-4s, .box-3s, .box-2s, .box-1s" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.2em 0.2em",
                                                                      "font-weight" = "400",
                                                                      "font-size" = "100%",
                                                                      "display" = "inline",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans",
                                                                    "border-radius" = "15px"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz",
                                                                    "border-radius" = "15px"),
   ".medium" = list("font-size" = "35px"),
   
   ".medium30" = list("font-size" = "30px"),
   
   ".jost-normal" = list("font-family" = "Jost",
                         "font-size" = "31px"),
   
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "60%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "34%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")


  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```
```{css, echo = FALSE}

.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}

.tiny .remark-code { /*Change made here*/
  font-size: 90% !important;
}
```

```{r setup2, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(showtext)
library(xaringanExtra)

xaringanExtra::use_scribble()

htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

```{r fonts, message=FALSE, echo=FALSE, warning=FALSE}
font.add.google("Fira Sans Condensed", "Fira Sans Condensed")
font.add.google("Fira Sans", "Fira Sans")
library(tidyverse)
```

<br>
<br>
.box-3Trans[Heteroskedasticity<br>
.small[(*hetero-skedasis*)]
]

--

.box-4Trans[*Different Dispersion*]

---
<br>


$$\log(Wages) = \beta_0 + \beta_1 Experience + \varepsilon$$

```{r wages_scatter, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}
library(tidyverse)
library(firasans)
library(hrbrthemes)
library(AER)
data("CPS1985")

CPS1985 <- CPS1985 %>% mutate(log_wage = log(wage))

ggplot(data = CPS1985, aes(y = log_wage, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
  theme_bw()+
  xlab("Experience (Years)") + ylab("log(Wage)") +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
<br>

$$\log(Wages) = \beta_0 + \beta_1 Experience + \varepsilon$$

Look at residuals, $Y_i - \hat{Y}_i$, where $\hat{Y}_i = \hat{\beta}_0 + \hat{\beta}_1Education_i$ are the fitted values.

```{r wages_residuals, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}
lm_wages <- lm(log_wage ~ experience, data = CPS1985)

CPS1985$residuals <- CPS1985$log_wage - predict(lm_wages)
CPS1985$fitted <- predict(lm_wages)

ggplot(data = CPS1985, aes(y = residuals, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
  theme_bw()+
  xlab("Experience (Years)") + ylab(expression(Y[i] - hat(Y)[i])) +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
<br>

$$\log(Wages) = \beta_0 + \beta_1 Experience + \varepsilon$$

Look at residuals, $Y_i - \hat{Y}_i$, where $\hat{Y}_i = \hat{\beta}_0 + \hat{\beta}_1Education_i$ are the fitted values.

```{r wages_residuals_pt2, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}
lm_wages <- lm(log_wage ~ experience, data = CPS1985)

CPS1985$residuals <- CPS1985$log_wage - predict(lm_wages)
CPS1985$fitted <- predict(lm_wages)

ggplot(data = CPS1985, aes(y = residuals, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
  geom_vline(aes(xintercept = 10), color = "darkgrey", lty = 2, lwd = 1.1) +
  geom_vline(aes(xintercept = 30), color = "darkgrey", lty = 2, lwd = 1.1) +
  theme_bw()+
  xlab("Experience (Years)") + ylab(expression(Y[i] - hat(Y)[i])) +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```


---

<br>
<br>
<br>
<br>
<br>
<br>

.box-5Trans[What happens in a Linear Probability Model?]

---
<br>

$$I(wage > 9) = \beta_0 + \beta_1 Experience + \varepsilon$$
--

```{r wages_scatter2, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}
CPS1985 <- CPS1985 %>% mutate(wage_over_9 = ifelse(wage>9, 1, 0))

ggplot(data = CPS1985, aes(y = wage_over_9, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
  theme_bw()+
  xlab("Experience (Years)") + ylab("Wage > $9") +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```


---
<br>

$$I(wage > 9) = \beta_0 + \beta_1 Experience + \varepsilon$$

- Remember that the variance of a binary variable $W$ is $Var(W)=p(1-p)$

```{r wages_residuals2, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}

lm_wages2 <- lm(wage_over_9 ~ experience, data = CPS1985)

CPS1985$residuals2 <- CPS1985$wage_over_9 - predict(lm_wages2)
CPS1985$fitted2 <- predict(lm_wages2)

ggplot(data = CPS1985, aes(y = residuals2, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
  theme_bw()+
  xlab("Experience (Years)") + ylab(expression(Y[i] - bar(Y))) +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
<br>

$$I(wage > 9) = \beta_0 + \beta_1 Experience + \varepsilon$$

- Remember that the variance of a binary variable $W$ is $Var(W)=p(1-p)$

```{r wages_residuals2_pt2, fig.height=5, fig.width=7.5, fig.align='center', dev='svg', echo=FALSE, warning=FALSE}

lm_wages2 <- lm(wage_over_9 ~ experience, data = CPS1985)

CPS1985$residuals2 <- CPS1985$wage_over_9 - predict(lm_wages2)
CPS1985$fitted2 <- predict(lm_wages2)

ggplot(data = CPS1985, aes(y = residuals2, x = experience)) +
  geom_point(fill = "white", color = "#900DA4", size = 3, pch = 21) +
  geom_smooth(method = "lm", se = FALSE, color = "#0D0887") +
    geom_vline(aes(xintercept = 0), color = "darkgrey", lty = 2, lwd = 1.1) +
  geom_vline(aes(xintercept = 40), color = "darkgrey", lty = 2, lwd = 1.1) +
  theme_bw()+
  xlab("Experience (Years)") + ylab(expression(Y[i] - bar(Y))) +
  theme_ipsum_fsc(plot_title_face = "bold",plot_title_size = 24) + #plain 
  
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=18),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=18),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_text(size=14),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```


---
# Why use robust standar errors?

Under heteroskedasticity, OLS is no longer **.darkorange[BLUE]**
  
  - **BLUE**: B(est) L(inear) U(nbiased) E(stimator)
  
We lose the **.darkorange[BEST]** part!

--

.pull-left[
- Under Homoskedasticity, $\varepsilon_i \sim N(0,\sigma^2)$:

  $$Var(\hat{\beta}_1) = \frac{\sum_i(X_i - \bar{X})^2\sigma^2}{(\sum_i(X_i - \bar{X})^2)^2} = \frac{\sigma^2}{\sum_i(X_i - \bar{X})^2} $$
]

--

.pull-right[
- Under Heteroskedasticity, $\varepsilon_i \sim N(0,\sigma_i^2)$:

  $$Var(\hat{\beta}_1) = \frac{\sum_i(X_i - \bar{X})^2\sigma_{i}}{(\sum_i(X_i - \bar{X})^2)^2}$$]

--

- Use **.darkorange[robust standard errors]** (i.e. Hubert-White SE) to take into account heteroskedasticity:

  - Use `lm_robust()` from the `estimatr` package.