---
title: "STA 235 - Multivariate Regression"
subtitle: "Spring 2021"
author: "McCombs School of Business, UT Austin"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Yanone Kaffeesatz",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Roboto", "300", "300i","400","500"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "130%",
  colors = c(
    red = "#f34213",
    purple = "#900DA4",
    darkpurple = "#61077a",
    darkorange = "#db5f12",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "none"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".large" = list("font-size" = "150%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Roboto",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".medium" = list("font-size" = "1.4em"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25"),
   
   ".box-1t,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.4)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz"),
    ".box-2t,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.4)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-3t,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.4)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-4t,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.4)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-5t,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.4)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-6t,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.4)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-7t,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.4)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "67%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "27%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")


  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25
```

# Remembering Regressions

- Linear Regression is a **.darkorange[very useful tool]**.

  - Simple supervised learning approach.
  - Many fancy methods are generalizations or extensions of linear regression!
  

.source[*Shoutout: Slides based on Andrew Heiss' "Program Evaluation for Public Service" course.*]


---
# Essential Parts of a Regression

.pull-left[
.box-2t.medium.sp-after-half[Y]
.box-2.medium.sp-after-half[Outcome Variable]
.box-2.medium.sp-after-half[Response Variable]
.box-2.medium.sp-after-half[Dependent Variable]

.box-2t.medium.sp-after-half[*Thing you want to explain or predict*]
  
]
--
.pull-right[
.box-2t.medium.sp-after-half[X]
.box-2.medium.sp-after-half[Explanatory Variable]
.box-2.medium.sp-after-half[Predictor Variable]
.box-2.medium.sp-after-half[Independent Variable]

.box-2t.medium.sp-after-half[*Thing you use to explain or predict Y*]
]

---
# Identify the variables

.pull-left[
.box-2.medium.sp-after-half[A study examines the effect of smoking on lung cancer]
]

---
# Identify the variables

.pull-left[
.box-2.medium.sp-after-half[A study examines the effect of smoking on lung cancer]
.box-4.medium.sp-after-half[Fantasy football fanatics predict the performance of a player based on past performance, health status, and characteristics of the opposite team]
]

---
# Identify the variables

.pull-left[
.box-2.medium.sp-after-half[A study examines the effect of smoking on lung cancer]
.box-4.medium.sp-after-half[Fantasy football fanatics predict the performance of a player based on past performance, health status, and characteristics of the opposite team]
]
.pull-right[
.box-5.medium.sp-after-half[You want to see if taking more AP classes in high school improves college grades]
]

---
# Identify the variables

.pull-left[
.box-2.medium.sp-after-half[A study examines the effect of smoking on lung cancer]
.box-4.medium.sp-after-half[Fantasy football fanatics predict the performance of a player based on past performance, health status, and characteristics of the opposite team]
]
.pull-right[
.box-5.medium.sp-after-half[You want to see if taking more AP classes in high school improves college grades]
.box-7.medium.sp-after-half[Netflix uses your past viewing history, the day of the week, and the time of the day to guess which show you want to watch next]
]
---
# Two Purposes of Regression

.pull-left[
.box-3t.medium.sp-after-half[Prediction]
.box-3.medium.sp-after-half[Forecast the future]
.box-3.medium.sp-after-half[Focus is on <b>Y</b>]

.box-3t.medium.sp-after-half[Netflix trying to guess your next show]
]
--
.pull-right[
.box-6t.medium.sp-after-half[Explanation]
.box-6.medium.sp-after-half[Explain the effect of <b>X</b> on <b>Y</b>]
.box-6.medium.sp-after-half[Focus is on <b>X</b>]

.box-6t.medium.sp-after-half[Netflix looking at the effect of time of the day on show selection]
]
---
# Let's write down regressions

- Focus on the **.darkorange[Netflix]** example:

.box-5.medium.sp-after-half[$$Y = \alpha + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \varepsilon$$]
---
background-position: 50% 50%
class: left, bottom, inverse
.big[
Let's review the syllabus
]
---
# References

- James, G. et al. (2017). "Introduction to Statistical Learning with Applications in R". *Chapter 3*

- Heiss, A. (2020). "Course: Program Evaluation for Public Service". *Slides for Regression and Inference*.

