---
title: "STA 235 - Causal Inference: Regression Discontinuity Design"
subtitle: "Spring 2021"
author: "McCombs School of Business, UT Austin"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "custom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","cols_macro.js"]
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.showtext = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

theme_xaringan(
  text_color = "#333f48",
  background_color = "#FFFFFF",
  accent_color = "#900DA4",
  text_font = "Fira Mono",
  text_font_use_google = TRUE,
  title_font = "Yanone Kaffeesatz",
  title_font_use_google = TRUE
)

style_mono_accent(
  #base_color = "#bf5700",
  extra_fonts = list(google_font("Fira Sans","200","300","400","500","600"),
                     google_font("Fira Sans Condensed","200","300","400","500","600")),
  base_color = "#333f48",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Roboto", "300", "300i","400","500"),
  code_font_google   = google_font("Fira Mono"),
  text_bold_color = "#333f48",
  text_font_size = "125%",
  colors = c(
    lightgrey = "#C0C0C0",
    red = "#f34213",
    purple = "#900DA4",
    darkpurple = "#61077a",
    darkorange = "#db5f12",
    orange = "#ff8811",
    green = "#136f63",
    white = "#FFFFFF"),
  extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "none"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                     "font-family" = "Yanone Kaffeesatz",
                     "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".tiny" = list("font-size" = "50%"),
    ".large" = list("font-size" = "150%"),
    ".source" = list("color" = "#8c8c8c",
                     "font-size" = "80%"),
    ".remark-slide table td#highlight" = list("background-color" = "#eee1f0",
                                  "color" = "#900DA4",
                                  "font-weight" = "500"),
   # ".remark-slide table thead th" = list(),
    ".title-slide h1" = list("font-weight" = "500"),
    ".title-slide h2" = list("font-weight" = "400",
                             "font-size" =  "170%"),
    ".title-slide h3" = list("font-family" = "Roboto",
                             "font-size" = "100%",
                             "font-weight" = "200"),
    ".center2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "50%",
                      "left" = "50%",
                      "-ms-transform" = "translate(-50%, -50%)",
                      "transform" = "translate(-50%, -50%)"),
   ".bottom2" = list("margin" = "0",
                      "position" = "absolute",
                      "top" = "90%",
                      "left" = "10%",
                      "-ms-transform" = "translate(-10%, -90%)",
                      "transform" = "translate(-10%, -90%)"),
    ".section-title h1" = list("color" = "#FFFFFF",
                               "font-size" = "2.3em",
                               "line-height" = "1.3"),
    ".medium" = list("font-size" = "1.4em"),
    ".sp-after-half" = list("margin-bottom" = "0.7em !important"),
    ".box-1,.box-1a,.box-1s,.box-1b,.box-1l,.box-1LA,.section-title-1" = list("background-color" = "#0D0887"),
    ".box-2,.box-2a,.box-2s,.box-2b,.box-2l,.box-2LA,.section-title-2" = list("background-color" = "#5601A4"),
    ".box-3,.box-3a,.box-3s,.box-3b,.box-3l,.box-3LA,.section-title-3" = list("background-color" = "#900DA4"),
    ".box-4,.box-4a,.box-4s,.box-4b,.box-4l,.box-4LA,.section-title-4" = list("background-color" = "#BF3984"),
    ".box-5,.box-5a,.box-5s,.box-5b,.box-5l,.box-5LA,.section-title-5" = list("background-color" = "#E16462"),
    ".box-6,.box-6a,.box-6s,.box-6b,.box-6l,.box-6LA,.section-title-6" = list("background-color" = "#F89441"),
    ".box-7,.box-7a,.box-7s,.box-7b,.box-7l,.box-7LA,.section-title-7" = list("background-color" = "#FCCE25"),
   
   ".box-1t,.box-1tL,.section-title-1t" = list("background-color" = "rgba(13, 8, 135,0.3)",
                                      "color"="rgba(13, 8, 135,1)",
                                      "font-family" = "Yanone Kaffeesatz"),
    ".box-2t,.box-2tL,.section-title-2t" = list("background-color" = "rgba(86, 1, 164,0.3)",
                                       "color" = "rgba(86, 1, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-3t,.box-3tL,.section-title-3t" = list("background-color" = "rgba(144, 13, 164,0.3)",
                                       "color" = "rgba(144, 13, 164,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-4t,.box-4tL,.section-title-4t" = list("background-color" = "rgba(191, 57, 132,0.3)",
                                       "color" = "rgba(191, 57, 132,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-5t,.box-5tL,.section-title-5t" = list("background-color" = "rgba(225, 100, 98,0.3)",
                                       "color" = "rgba(225, 100, 98,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-6t,.box-6tL,.section-title-6t" = list("background-color" = "rgba(248, 148, 65,0.3)",
                                       "color" = "rgba(248, 148, 65,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
    ".box-7t,.box-7tL,.section-title-7t" = list("background-color" = "rgba(252, 206, 37,0.3)",
                                       "color" = "rgba(252, 206, 37,1)",
                                       "font-family" = "Yanone Kaffeesatz"),
   
   ".box-7t, .box-6t, .box-5t, .box-4t, .box-3t, .box-2t, .box-1t" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center"),
   
   ".box-7tL, .box-6tL, .box-5tL, .box-4tL, .box-3tL, .box-2tL, .box-1tL" = list("margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "50px",
                                                                    "display" = "table",
                                                                    "text-align" = "center"),
   
    ".box-7, .box-6, .box-5, .box-4, .box-3, .box-2, .box-1" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "31px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                    "font-family" = "Fira Sans"),
    ".box-7a, .box-6a, .box-5a, .box-4a, .box-3a, .box-2a, .box-1a" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans"),
   ".box-7s, .box-6s, .box-5s, .box-4s, .box-3s, .box-2s, .box-1s" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.2em 0.2em",
                                                                      "font-weight" = "400",
                                                                      "font-size" = "100%",
                                                                      "display" = "inline",
                                                                      "text-align" = "center",
                                                                      "font-family" = "Fira Sans"),
       ".box-7b, .box-6b, .box-5b, .box-4b, .box-3b, .box-2b, .box-1b" = list("color" = "#FFFFFF",
                                                                          "left" = "0px",
                                                                          "overflow" = "hidden",
                                                                      "padding" = "0.4em 0.4em",
                                                                      "font-weight" = "600",
                                                                      "font-size" = "25px",
                                                                      "display" = "table",
                                                                      "text-align" = "left",
                                                                      "font-family" = "Fira Sans"),
   ".box-7l, .box-6l, .box-5l, .box-4l, .box-3l, .box-2l, .box-1l" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "45px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".box-7LA, .box-6LA, .box-5LA, .box-4LA, .box-3LA, .box-2LA, .box-1LA" = list("color" = "#FFFFFF",
                                                                    "margin" = "0em auto",
                                                                    "overflow" = "hidden",
                                                                    "padding" = "0.4em 0.4em",
                                                                    "font-weight" = "600",
                                                                    "font-size" = "55px",
                                                                    "display" = "table",
                                                                    "text-align" = "center",
                                                                      "font-family" = "Yanone Kaffeesatz"),
   ".image-80 img" = list("scale" = "80%"),
   ".pull-left-little_l" = list("float" = "left",
                                "width" = "67%"),
   ".pull-right-little_l" = list("float" = "right",
                                "width" = "27%"),
   ".pull-left-little_r" = list("float" = "left",
                                "width" = "27%"),
   ".pull-right-little_r" = list("float" = "right",
                                "width" = "67%")


  )
)

#,"li" = list("font-size" = "150%")
#    "li" = list("font-size" = "110%"),
#    "ul" = list("font-size" = "110%"),
#color palette
#5601A4
#900DA4
#F89441
#FCCE25

knitr::opts_chunk$set(message = FALSE)

```
```{css, echo = FALSE}

.small .remark-code { /*Change made here*/
  font-size: 80% !important;
}

.tiny .remark-code { /*Change made here*/
  font-size: 90% !important;
}
```

```{r setup2, echo=FALSE, message=FALSE}
library(knitr)
library(showtext)
```

```{r fonts, message=FALSE, echo=FALSE}
font.add.google("Fira Sans Condensed", "Fira Sans Condensed")
font.add.google("Fira Sans", "Fira Sans")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gganimate)
library(ggthemes)
library(gifski)
library(plotly)
library(leaflet)
library(ggplot2)
library(firasans)
library(hrbrthemes)
library(extrafont)
library("RColorBrewer")
library(viridis)
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
library("firasans")
library(hrbrthemes)
library(extrafont)


```


# Another identification strategy

- We have seen:

.box-2[RCTs]

.box-3[Selection on observables]

.box-4[Natural experiments]

.box-6[Differences-in-Differences]

--

.box-7LA[Regression Discontinuity Designs]


---
background-position: 50% 50%
class: left, bottom, inverse
.big[
I'm on the edge [of glory?]
]
---
# Introduction to Regression Discontinuity Designs


<br>
<br>
<br>
.box-2LA[Regression Discontinuity (RD) Designs]
--
<br>
<br>
.box-2tL[Arbitrary rules determine treatment assignment]

--
<br>
<br>
.box-3t[E.g.: If you are above a threshold, you are assigned to treatment, and if your below, you are not (or vice versa)]

---
# Key Terms

.box-4LA[Running/ forcing variable]

.box-4t[Index or measure that determines eligibility]

--

.box-7LA[Cutoff/ cutpoint/ threshold]

.box-7t[Number that formally assigns you to a program or treatment]

---
# Hypothetical tutoring program

--

.box-5[Students take an entrance exam]

--

.box-5tL[Those who score 70 or lower<br>get a free tutor for the year]

--

.box-5[Students then take an exit exam<br>at the end of the year]

---
# Assignment based on entrance score

```{r tutoring-running, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(kableExtra)
library(scales)
library(patchwork)
library(latex2exp)
library(rdrobust)
library(rddensity)

# Fake program data
set.seed(1234)
num_students <- 1000
tutoring <- tibble(
  id = 1:num_students,
  entrance_exam = rbeta(num_students, shape1 = 7, shape2 = 2),
  exit_exam = rbeta(num_students, shape1 = 5, shape2 = 3)
) %>% 
  mutate(entrance_exam = entrance_exam * 100,
         tutoring = entrance_exam <= 70) %>% 
  mutate(exit_exam = exit_exam * 40 + 10 * tutoring + entrance_exam / 2) %>% 
  mutate(tutoring_fuzzy = ifelse(entrance_exam > 60 & entrance_exam < 80,
                                 sample(c(TRUE, FALSE), n(), replace = TRUE),
                                 tutoring)) %>% 
  mutate(tutoring_text = factor(tutoring, levels = c(FALSE, TRUE),
                                labels = c("No tutor", "Tutor")),
         tutoring_fuzzy_text = factor(tutoring_fuzzy, levels = c(FALSE, TRUE),
                                      labels = c("No tutor", "Tutor"))) %>% 
  mutate(entrance_centered = entrance_exam - 70)

ggplot(tutoring, aes(x = entrance_exam, y = tutoring_text, fill = tutoring_text)) +
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_point(size = 5, pch = 21, color = "white", alpha = 0.7,
             position = position_jitter(width = 0, height = 0.2, seed = 1234)) + 
  labs(x = "Entrance exam score", y = NULL) + 
  guides(fill = FALSE) +
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```

---
# Let's look at the area close to the cutoff

```{r tutoring-running-threshold, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}

ggplot(tutoring, aes(x = entrance_exam, y = tutoring_text, fill = tutoring_text)) +
  ggplot2::annotate(geom = "rect", fill = "grey50", alpha = 0.25, ymin = -Inf, ymax = Inf,
           xmin = 70 - 5,  xmax = 70 + 5) +
  ggplot2::annotate(geom = "rect", fill = "grey50", alpha = 0.5, ymin = -Inf, ymax = Inf,
           xmin = 70 - 2,  xmax = 70 + 2) +
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_point(size = 5, pch = 21, color = "white", alpha = 0.7,
             position = position_jitter(width = 0, height = 0.2, seed = 1234)) + 
  labs(x = "Entrance exam score", y = NULL) + 
  guides(fill = FALSE) +
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```

---
# Let's get closer

```{r tutoring-running-threshold-zoomed, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}

ggplot(tutoring, aes(x = entrance_exam, y = tutoring_text, fill = tutoring_text)) +
  ggplot2::annotate(geom = "rect", fill = "grey50", alpha = 0.25, ymin = -Inf, ymax = Inf,
           xmin = 70 - 5,  xmax = 70 + 5) +
  ggplot2::annotate(geom = "rect", fill = "grey50", alpha = 0.5, ymin = -Inf, ymax = Inf,
           xmin = 70 - 2,  xmax = 70 + 2) +
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_point(size = 5, pch = 21, color = "white", alpha = 0.7,
             position = position_jitter(width = 0, height = 0.2, seed = 1234)) + 
  labs(x = "Entrance exam score", y = NULL) + 
  guides(fill = FALSE) +
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  coord_cartesian(xlim = c(70 - 6, 70 + 6)) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```
---
# Causal inference intuition

.box-2LA[Observations right before and after the threshold are essentially the same]

--

.box-4LA[Pseudo treatment and control groups!]

--

.box-7LA[Compare outcomes right at the cutoff]

---
# Exit exam results according to running variable

```{r tutoring-outcome, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 1) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  labs(x = "Entrance exam score", y = "Exit exam score") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```

---
# Fit a regression at the right and left side of the cutoff

```{r tutoring-outcome-lines, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "lm", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "lm", size = 2, show.legend = FALSE) +
  labs(x = "Entrance exam score", y = "Exit exam score") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```

---
# Fit a regression at the right and left side of the cutoff

```{r tutoring-outcome-edlta, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
rdd_tutoring <- lm(exit_exam ~ entrance_centered + tutoring, data = tutoring) %>% 
  tidy()
effect_control <- filter(rdd_tutoring, term == "(Intercept)")$estimate
late <- filter(rdd_tutoring, term == "tutoringTRUE")$estimate
effect_treatment <- effect_control + late
ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "lm", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "lm", size = 2, show.legend = FALSE) +
  ggplot2::annotate(geom = "segment", x = 70, xend = 70, 
           y = effect_control, yend = effect_treatment,
           size = 4, color = "#5601A4") +
  ggplot2::annotate(geom = "label", x = 82, y = effect_treatment - (late / 2),
           label = latex2exp("$\\delta$\n(Causal effect)"), family = "Fira Sans Condensed", fontface = "bold",
           color = "white", fill = "#5601A4", size = 7, label.padding = unit(0.75, "lines")) +
  labs(x = "Entrance exam score", y = "Exit exam score") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed")
```

---

.center2[
.box-5LA[You can find discontinuities everywhere!]
]

---
# Geographic discontinuities


---
# Time discontinuities


---
# Voting discontinuities

---
background-position: 50% 50%
class: left, bottom, inverse
.big[
How do we do RDs in practice?
]

---
# Behind the scenes of RDs

- Basically, regression discontinuities work under an **.darkorange[asymptotic assumption]**:

--

- Let $Y_i$ be the outcome of interest, $Z_i$ the treatment assignment, $R_i$ the running variable, and $c$ the cutoff score:

$$Z_i =   \left\{
\begin{array}{ll}
      0 & R_i\leq c \\
      1 & R_i>c \\
\end{array} 
\right.$$

--

- Then, we can define the treatment effect $\delta$ as:

$$\delta = \lim_{\epsilon \rightarrow 0^+}E[Y_i| R_i = c + \epsilon] - \lim_{\epsilon \rightarrow 0^-}E[Y_i| R_i = c + \epsilon]$$
---
# What does the limit expression mean?

```{r tutoring-limit, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
tutoring$exit_exam[tutoring$entrance_exam>70] = tutoring$exit_exam[tutoring$entrance_exam>70] + 25

ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "loess", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "loess", size = 2, show.legend = FALSE) +
  labs(x = "Running Variable", y = "Outcome") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed") +
  theme_ipsum() +
  #guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
        #axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```
---
# What does the limit expression mean?

```{r tutoring-limit2, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}

ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "loess", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "loess", size = 2, show.legend = FALSE) +
  labs(x = "Running Variable", y = "Outcome") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed") +
  theme_ipsum() +
  geom_segment(aes(x=60,y=70,xend=70,yend=70),lwd=0.8,lty=1,
               arrow=arrow(length = unit(0.03, "npc"),type="closed"),arrow.fill="dark grey",col="dark grey") +
  #guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
        #axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
# What does the limit expression mean?

```{r tutoring-limit3, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}

ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "loess", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "loess", size = 2, show.legend = FALSE) +
  labs(x = "Running Variable", y = "Outcome") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed") +
  theme_ipsum() +
  geom_segment(aes(x=80,y=85,xend=70,yend=85),lwd=0.8,lty=1,
               arrow=arrow(length = unit(0.03, "npc"),type="closed"),arrow.fill="dark grey",col="dark grey") +
  geom_segment(aes(x=60,y=70,xend=70,yend=70),lwd=0.8,lty=1,
               arrow=arrow(length = unit(0.03, "npc"),type="closed"),arrow.fill="dark grey",col="dark grey") +
  #guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
        #axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
.center2[
.box-2LA[What is the estimand we are estimating?]
]

---
.center2[
.box-2LA[Local Average Treatment Effect (LATE) for units at R=c]
]
---
# Conditions required for identificaton

- Threshold rule **.darkorange[exists]** and cutoff point is **.darkorange[known]**

--

- The running variable $R_i$ is **.darkorange[continuous]** near $c$.

--

- Key assumption:

.box-5LA[Continuity of E[Y(1)|R] and E[Y(0)|R] at R=c]

---
# Potential outcomes need to be smooth across the threshold

```{r po-smooth, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
# Fake program data
set.seed(1234)
num_students <- 1000
tutoring <- tibble(
  id = 1:num_students,
  entrance_exam = rbeta(num_students, shape1 = 7, shape2 = 2),
  exit_exam = rbeta(num_students, shape1 = 5, shape2 = 3)
) %>% 
  mutate(entrance_exam = entrance_exam * 100,
         tutoring = entrance_exam <= 70) %>% 
  mutate(po = exit_exam * 40 + entrance_exam / 2,
    exit_exam = exit_exam * 40 + 10 * tutoring + entrance_exam / 2) %>% 
  mutate(tutoring_fuzzy = ifelse(entrance_exam > 60 & entrance_exam < 80,
                                 sample(c(TRUE, FALSE), n(), replace = TRUE),
                                 tutoring)) %>% 
  mutate(tutoring_text = factor(tutoring, levels = c(FALSE, TRUE),
                                labels = c("No tutor", "Tutor")),
         tutoring_fuzzy_text = factor(tutoring_fuzzy, levels = c(FALSE, TRUE),
                                      labels = c("No tutor", "Tutor"))) %>% 
  mutate(entrance_centered = entrance_exam - 70)

tutoring$exit_exam[tutoring$entrance_exam>70] = tutoring$exit_exam[tutoring$entrance_exam>70] + 25
#tutoring$po = tutoring$po + 25

ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "loess", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "loess", size = 2, show.legend = FALSE) +
  
  geom_smooth(data = filter(tutoring, entrance_exam <= 70), aes(x = entrance_exam, y = po + 25),
              method = "loess", se = FALSE, lty = 2,size = 2, show.legend = FALSE, color = "#900DA4") +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),aes(x = entrance_exam, y = po + 10),
              method = "loess", size = 2, se = FALSE, lty = 2, show.legend = FALSE, color = "#F89441") +
  
  labs(x = "Running Variable", y = "Outcome") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed") +
  theme_ipsum() +
  #guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
        #axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---
# Potential outcomes need to be smooth across the threshold

```{r po-not-smooth2, fig.height=5.25, fig.width=12, fig.align='center', dev='svg', echo=FALSE, warning=FALSE, message = FALSE}
# Fake program data
set.seed(1234)
num_students <- 1000
tutoring <- tibble(
  id = 1:num_students,
  entrance_exam = rbeta(num_students, shape1 = 7, shape2 = 2),
  exit_exam = rbeta(num_students, shape1 = 5, shape2 = 3)
) %>% 
  mutate(entrance_exam = entrance_exam * 100,
         tutoring = entrance_exam <= 70) %>% 
  mutate(po = exit_exam * 40 + entrance_exam / 2,
    exit_exam = exit_exam * 40 + 10 * tutoring + entrance_exam / 2) %>% 
  mutate(tutoring_fuzzy = ifelse(entrance_exam > 60 & entrance_exam < 80,
                                 sample(c(TRUE, FALSE), n(), replace = TRUE),
                                 tutoring)) %>% 
  mutate(tutoring_text = factor(tutoring, levels = c(FALSE, TRUE),
                                labels = c("No tutor", "Tutor")),
         tutoring_fuzzy_text = factor(tutoring_fuzzy, levels = c(FALSE, TRUE),
                                      labels = c("No tutor", "Tutor"))) %>% 
  mutate(entrance_centered = entrance_exam - 70)

tutoring$exit_exam[tutoring$entrance_exam>70] = tutoring$exit_exam[tutoring$entrance_exam>70] + 25
#tutoring$po = tutoring$po + 25

ggplot(tutoring, aes(x = entrance_exam, y = exit_exam, fill = tutoring_text, color = tutoring_text)) +
  geom_point(size = 4, pch = 21, color = "white", alpha = 0.25) + 
  geom_vline(xintercept = 70, size = 2, color = "#FFDC00") + 
  geom_smooth(data = filter(tutoring, entrance_exam <= 70),
              method = "loess", size = 2, show.legend = FALSE) +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),
              method = "loess", size = 2, show.legend = FALSE) +
  
  geom_smooth(data = filter(tutoring, entrance_exam <= 70), aes(x = entrance_exam, y = po + 15),
              method = "loess", se = FALSE, lty = 2,size = 2, show.legend = FALSE, color = "#900DA4") +
  geom_smooth(data = filter(tutoring, entrance_exam > 70),aes(x = entrance_exam, y = po + 10),
              method = "loess", size = 2, se = FALSE, lty = 2, show.legend = FALSE, color = "#F89441") +
  
  labs(x = "Running Variable", y = "Outcome") + 
  scale_fill_manual(values = c("#900DA4", "#F89441"), name = NULL) +
  scale_color_manual(values = c("#900DA4", "#F89441")) +
  theme_bw(base_size = 28, base_family = "Fira Sans Condensed") +
  theme_ipsum() +
  #guides(fill = guide_legend(reverse = TRUE, override.aes = list(alpha = 1)), color = FALSE) +
  theme(plot.margin=unit(c(1,1,1.5,1.2),"cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
        #axis.line = element_line(colour = "dark grey"))+
  theme(axis.title.x = element_text(size=16),#margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size=16),#margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.text.y = element_blank(),legend.position="none",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.background = element_rect(fill="white",colour ="white"),
        title = element_text(size=14))
```

---

.center2[
.box-2LA[Can you think situations where that could happen?]
]

---
# Work in groups

<br>
.box-3[1) Each group will be given a task and some code]

--
<br>
<br>

.box-5[2) You need to complete the code and discuss the results]
<!-- pagedown::chrome_print('C:/Users/mc72574/Dropbox/Hugo/Sites/sta235/exampleSite/content/Classes/Week7/1_DD/sp2021_sta235_9_d_cont.html') -->