---
title: "Homework 2: Causal Inference (Part 1)"
author: "STA 235"
date: "Due on March 1st, 12:00pm"
output: 
  html_document:
    css: style.css 
    toc: no
  pdf_document: 
    latex_engine: xelatex
    css: style.css
    toc: no
---

# Task 1: Let's run an experiment

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(modelsummary)

# Load insurance data
insurance <- read.csv("https://raw.githubusercontent.com/maibennett/sta235/main/exampleSite/content/Assignments/Homework1/data/insurance.csv")
```

We have a dataset of over 1,300 people with their insurance charges and socio-demographic characteristics.

- `age`: The insurer's age.
- `sex`: Whether the insurer is male or female.
- `bmi`: A measure of Body Mass Index, which is constructed by dividing the person's weight (in kg.) over their squared height (in cm.)
- `children`: Number of children. 
- `smoker`: Whether the person is a smoker or not.
- `region`: Region in the US in which they reside.
- `charges`: Insurance charges.

---

## Exploratory analysis

What is the relationship between charges and age? This plot shows some initial trends:

```{r plot-age-charges}
ggplot(data = insurance, 
       aes(x = age, y = charges)) +
  geom_point(color = "dark orange", fill = alpha("dark orange", 0.5), size = 3, pch = 21) +
  theme_bw()
  
```


### Question 1.1: 
**Describe the relationship between age and charges based on the previous plot.**

### Question 1.2: 
**Make a new plot that colors these points by whether a person is a smoker or not. What can you tell about the relationship between age and charges?**

### Question 1.3: 
**Now, show a new plot with the relationship between BMI and charges. What can you say about that relationship? Does the relationship look linear?**

---

## Let's build some models

### Question 1.4: 
**Using the previous information, build a regression using age, body mass index, smoker category, and sex as covariates, and charges as the dependent variable. Use any variable transformation that you believe appropriate (and justify it). Interpret the coefficients for this model.**



Finally, let's add region to the previous regression. 

### Question 1.5: 
**If we wanted to predict an individual's insurance charges, which model do you think would be better and why? Show your results.** 


---

# Task 2: Real-world data

```{r setup2, warning=FALSE, message=FALSE}
library(estimatr)

nba <- read.csv("https://raw.githubusercontent.com/maibennett/sta235/main/exampleSite/content/Assignments/Homework1/data/nba.csv")
```

In this section, we will work with NBA data analyzing.

- `name`: Player's name.
- `GP`: Games played.
- `MIN`: Minutes played.
- `PTS`: Points per game. 
- `FGM`: Field goals made.
- `FGA`: Field goals attempts.
- `FGP`: Field goals percentage.
- `X3PM`: 3 points made.
- `X3PA`: 3 points attempts.
- `X3PP`: 3 points percentage.
- `FTM`: Free throw made.
- `FTA`: Free throw attempts.
- `FTP`: Free throw percentage.
- `OREB`: Offensive rebounds.
- `DREB`: Defensive rebounds.
- `REB`: Rebounds.
- `AST`: Assists.
- `STL`: Steals.
- `BLK`: Blocks.
- `TOV`: Turnovers.
- `TARGET_5Yrs`: Career length in NBA $\geq$ 5 years (1) or $<$ 5 years (0).

---
## Fitting some models

The following linear probability model regresses some rookie stats (games played, average points per game, 3-point percentage, rebounds, and assists) on whether a player is still on the NBA five years later.   

### Question 2.1: 
**Interpret the coefficient for games played and 3-point percentage in this setting**

```{r, echo=FALSE}
lm2.1 <- estimatr::lm_robust(TARGET_5Yrs ~ GP + PTS + X3PP + REB + AST, data = nba)

modelsummary(list(Target_5Yrs = lm2.1), stars = TRUE)
```


Now, fit the same model using a logistic regression.

### Question 2.2: 
**Interpret the same coefficients for this model (games played and 3-point percentage). How do they compare?**


---
## Predictions

Now imagine you want to make a prediction for two rookies in particular: Smith Roberts and Tyler Stevens. Their stats are given here:

```{r warning=FALSE, message=FALSE}

rookies <- data.frame(Name = c("Smith Roberts","Tyler Stevens"),
                      GP = c(91,15),
                      PTS = c(28,42),
                      X3PP = c(66.1,75.5),
                      REB = c(2,5.1),
                      AST = c(1.5,2))

```

### Question 2.3: 
**Compare the predictions given by the Linear Probability Model and the Logistic Regression. Which model would you use and why?**


---

# Task 3: Introduction to Causal Inference

Let's go back to the insurance data. Imagine that you now have also data for people that are uninsured:

```{r, echo=FALSE}
# Load all insurance data
insurance_all <- read.csv("https://raw.githubusercontent.com/maibennett/sta235/main/exampleSite/content/Assignments/Homework1/data/insurance_all.csv")

```

### Question 3.1: 
**What is the relationship between being insured and BMI? and what about smoking? Draw some figures! Comment on your results.**


We are interested in a problem called "Moral Hazard", which refers to the idea that people tend to take more risks after they get insurance compared to those that are uninsured or even compared to their previous behavior (e.g. drive faster, take worse care of their health, etc.). A colleague says that they know how to identify it, and shows you the following results:

```{r, echo=FALSE}
lm3.1 <- lm(bmi ~ insured + factor(smoker) + age + factor(sex) + factor(region), data = insurance_all)

modelsummary(list(BMI = lm3.1), stars = TRUE)
```

### Question 3.2: 
**Can we identify the effect of being insured on BMI with this model? Why or why not? Can you see a collider problem?**


---

## Design your own study:

### Question 3.3: 
**If you wanted to identify the causal effect of being insured on BMI, how would you do it? Would you need additional data?**


<!-- pagedown::chrome_print('C:/Users/mc72574/Dropbox/UT/Teaching/Assignments/Homework1/STA235_Homework1.html') -->
