---
title: "Homework 1: Regressions and Potential Outcomes"
author: "STA 235"
date: "Due on Feb. 15th, 12:00pm"
output: 
  html_document: 
    toc: no
  pdf_document: 
    latex_engine: xelatex
    toc: no
  word_document: 
    toc: no
---

# Task 1: How much should insurance cost?

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(modelsummary)

# Load insurance data
insurance <- read.csv("https://raw.githubusercontent.com/maibennett/sta235/main/exampleSite/content/Assignments/Homework1/data/insurance.csv")
```

We have a dataset of over 1,300 people with their insurance charges and socio-demographic characteristics.

- `age`: The insurer's age.
- `sex`: Whether the insurer is male or female.
- `bmi`: A measure of Body Mass Index, which is constructed by dividing the person's weight (in kg.) over their squared height (in cm.)
- `children`: Number of children. 
- `smoker`: Whether the person is a smoker or not.
- `region`: Region in the US in which they reside.
- `charges`: Insurance charges.

## Exploratory analysis

What is the relationship between charges and age? This plot shows some initial trends:

```{r plot-age-charges}
ggplot(data = insurance, 
       aes(x = age, y = charges)) +
  geom_point(color = "dark orange", fill = alpha("dark orange", 0.5), size = 3, pch = 21) +
  theme_bw()
  
```

**Question 1: Describe the relationship between age and charges based on the previous plot.**


**Question 2: Make a new plot that colors these points by whether a person is a smoker or not. What can you tell about the relationship between age and charges?**

```{r plot-age-charges-by-smoker}
ggplot(data = insurance, 
       aes(x = age, y = charges, fill = factor(smoker), color = factor(smoker))) +
  geom_point(size = 3, pch = 21) +
  theme_bw()
```


**Question 3: Now, show a new plot with the relationship between BMI and charges. What can you say about that relationship? Does the relationship look linear?**

```{r plot-bmi-charges}
ggplot(data = insurance, 
       aes(x = bmi, y = charges)) +
  geom_point(color = "purple", fill = alpha("purple", 0.5), size = 3, pch = 21) +
  theme_bw()
```

## Let's build some models

**Question 4: Using the previous information, build a regression using age, body mass index, smoker category, and sex as covariates, and charges as the dependent variable. Interpret the coefficients for this model.**
```{r, echo=FALSE}
#summary(lm(charges ~ age + bmi + I(bmi^2) + factor(smoker) + factor(sex), data = insurance))
lm1 <- lm(charges ~ age + bmi + I(bmi^2) + factor(smoker) + factor(sex), data = insurance)

modelsummary(list(lm1), stars = TRUE)
```


Finally, let's add region to the previous regression. 

**Question 5: If we wanted to predict an individual's charges, which model do you think would be better and why?** 
```{r, echo=FALSE}
#summary(lm(charges ~ age + bmi + I(bmi^2) + factor(smoker) + factor(sex) + factor(region), data = insurance))

lm2 <- lm(charges ~ age + bmi + I(bmi^2) + factor(smoker) + factor(sex) + factor(region), data = insurance)

modelsummary(list(lm1, lm2),
             coef_omit = "factor(region)", stars = TRUE)
```

#Task 2: Models with binary outcomes
